Do larger catchments respond different to forest cover change? Re-analysing a global data set.

### 2021-10-13 paper/Forest_and_Water/Forest_and_Water.Rmd
updates paper 14102021

In addition, additional variables added were the latitude and longitude for the center of the watershed as an approximation of its spatial location. Using this information  annual average reference evapotranspiration (E0) was extracted from **XXXXX** if a value of E0 was not available from the original papers. For large watersheds, this value, similar to annual average rainfall, is only an approximation of the climate at the location.

The length of the study can be a variable influencing the change in flow [e.g. @jackson2005] and therefore, the length, starting data and completion date of the different studies was extracted from the references provided by @zhang2017.

@@

@@

<!-- analysis -->
Create a new data frame in which all the change in forestry is positive, but an additional column is added that indicates whether it was a forest cover increase or decrease.

@@

@@

```
Change in forest cover has a significant impact on the change in flow, and the change in flow is different under deforestation compared to reforestation. In fact, reforestation has a 32% lower impact than deforestation. Overall model explanatory power is just `r summary(linear_model_simple)$r.squared`

@@

linear_model2 <- gam(DeltaQf_perc ~ DeltaF_perc_pos + Forest_Sign + Area_km2 + Pa_mm, data = Zhang_all2)
summary(linear_model2)
gam.check(linear_model2)
```
Area is not a significant predictor of the change in flow.
But decreases in forestry definitely have a different effect from increases in flow.

<!-- This shows that Area is not significant -->

<!-- include Lat and Long (later E0) -->

```{r model2a, echo = F}
linear_model_all_log <- gam(log10(DeltaQf_perc/100+1) ~ DeltaF_perc + Area_km2 + Pa_mm + Latitude + Longitude, data = Zhang_all)
summary(linear_model_all_log)
gam.check(linear_model_all_log)
```

>>>>

In addition, additional variables added were the latitude and longitude for the center of the watershed as an approximation of its spatial location. Using this information  annual average potential evapotranspiration (E_{pot}) was extracted from the MODIS16 ET product, if a value of E0 was not available from the original papers. This involved downloading the MODIS product for PET at 500 m scale for the approximate catchment centroid using the package MODISTools via the [MODIS/VIIRS subsets](https://modis.ornl.gov/data/modis_webservice.html) for the period 1 January 2000 - 31 December 2020. The average annual PET calculated from this series was used as the an approximation of average annual E0. For large watersheds, this value, similar to annual average rainfall, is only an approximation of the climate at the location.

The length of the study can be a variable influencing the change in flow [e.g. @jackson2005], as for example, more mature plantations are thought to have smaller impacts on flow. Therefore, the length of the study calculate as the difference between the starting data and completion date of the different studies was extracted from the references provided by @zhang2017.

@@

**This methods section is no longer correct and needs rewriting**

@@

<!-- analysis -->
#### description of the data

The overall dataset contains `r nrow(Zhang_all)` observations of changes in flow. The overall distribution of changes in flow is highly skewed as is the distribution of changes in forest cover and Area. The values of changes in flow greater than 100\% and smaller than -100\% clearly create long tails on the change in flow distribution. Note also the large number of studies with 100\% forest cover reduction.

```{r}
Zhang_all %>%
  pivot_longer(c(DeltaF_perc, DeltaQf_perc, Area_km2, Pa_mm), 
               names_to = "variable", values_to = "value") %>%
  ggplot(aes(value)) + geom_histogram(fill = "steelblue", bins=50) + theme_bw() + facet_wrap(~variable, scales = "free")
```

The changes in forest cover contain both positive (forestation) and negative values (deforestation). In @zhang 2017, these changes were analysed jointly, which assumes that the effect on the change in flow is linear and non-hysteretic. However, it is a reasonable hypothesis that this is not the case, and that the impact of an increase in forest cover is different from the same fractional decrease in forest cover. To be able to analyse this difference, all the change in forest cover is converted to positive values, but an additional column is added that indicates whether it was a forest cover increase or decrease.

@@

#### The initial relationship between change in forest cover and streamflow

@@

```
While the overall variance explained in this model is not high at `r round(summary(linear_model_simple)$r.sq,2)`, it clearly indicates the hypothesised relationship between the change in forest cover and the change in flow. The model suggests that for every 1\% change in forest cover, on the average, the flow changes 0.5\%. However the change in flow is different for forest cover decreases compared to forest cover increases. In fact, forest cover increases decrease flow by 32\% less than a similar decrease in forest cover causes flow to increase. So roughly speaking, a 1\% forest cover increase on the average decreases flow by $(1 - 0.32)*0.5\%$, while a the percentage forest cover decrease will increase flow by 0.5\%.

It is however clear from lack of explaining power, that there could be confounding factors, as alluded to in the methods. The obvious ones being catchment dryness and area (following @zhang2017).

<!-- At the moment using catchment rainfall, we probably want to straight away use P/E0, and explain that rainfall itself might not be a good predictor. Alternatively we can first show that Latitude is a predictor -->

@@

linear_model2 <- gam(DeltaQf_perc ~ DeltaF_perc_pos + Forest_Sign + Area_km2 + Pa_mm, data = Zhang_all2)
summary(linear_model2)
gam.check(linear_model2)
```

Including area and annual precipitation does not really improve the overall explaining power of the model, in fact, annual precipitation appears to be only a very small confounding factor, representing only a -0.01/% partial effect in the change in streamflow, holding all other factors constant. The catchment area does not appear to have an effect at all in contrast to earlier reported studies [@zhang2017; @filoso2017]. The main effects remain the change in forest cover and whether this is an increase or decrease

#### The effect of location on the globe
<!-- include Lat and Long (later E0) -->

```{r model2a, echo = F}
linear_model_all_log <- gam(DeltaQf_perc ~ DeltaF_perc + Forest_Sign + Area_km2 + Pa_mm + Latitude + Longitude, data = Zhang_all2)
summary(linear_model_all_log)
gam.check(linear_model_all_log)
```

### 2021-10-13 paper/Forest_and_Water/Forest_and_Water.tex
updates paper 14102021

In addition, additional variables added were the latitude and longitude
for the center of the watershed as an approximation of its spatial
location. Using this information annual average reference
evapotranspiration (E0) was extracted from \textbf{XXXXX} if a value of
E0 was not available from the original papers. For large watersheds,
this value, similar to annual average rainfall, is only an approximation
of the climate at the location.

The length of the study can be a variable influencing the change in flow
(e.g.~Jackson et al., 2005) and therefore, the length, starting data and
completion date of the different studies was extracted from the
references provided by Zhang et al. (2017).

@@

@@

\begin{verbatim}
## Loading required package: mgcv
\end{verbatim}

\begin{verbatim}
## Loading required package: nlme
\end{verbatim}

\begin{verbatim}
## 
## Attaching package: 'nlme'
\end{verbatim}

\begin{verbatim}
## The following object is masked from 'package:dplyr':
## 
##     collapse
\end{verbatim}

\begin{verbatim}
## This is mgcv 1.8-35. For overview type 'help("mgcv-package")'.
\end{verbatim}

@@

\begin{verbatim}
## 
## Family: gaussian 
## Link function: identity 
## 
## Formula:
## log10(DeltaQf_perc/100 + 1) ~ DeltaF_perc + Area_km2 + Pa_mm
## 
## Parametric coefficients:
##               Estimate Std. Error t value Pr(>|t|)    
## (Intercept)  4.430e-02  1.759e-02   2.519   0.0123 *  
## DeltaF_perc -1.345e-03  1.468e-04  -9.163   <2e-16 ***
## Area_km2    -1.657e-08  3.057e-08  -0.542   0.5883    
## Pa_mm       -7.989e-06  1.164e-05  -0.686   0.4930    
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## 
## R-sq.(adj) =   0.21   Deviance explained = 21.8%
## GCV = 0.018116  Scale est. = 0.017882  n = 310
\end{verbatim}

\begin{verbatim}
## 
## Method: GCV   Optimizer: magic
## Model required no smoothing parameter selectionModel rank =  4 / 4
\end{verbatim}

\begin{verbatim}
## 
## Family: gaussian 
## Link function: identity 
## 
## Formula:
## log10(DeltaQf_perc/100 + 1) ~ DeltaF_perc + Area_km2 + Pa_mm + 
##     Latitude + Longitude
## 
## Parametric coefficients:
##               Estimate Std. Error t value Pr(>|t|)    
## (Intercept)  7.296e-02  1.964e-02   3.715 0.000243 ***
## DeltaF_perc -1.552e-03  1.526e-04 -10.173  < 2e-16 ***
## Area_km2    -2.963e-08  2.996e-08  -0.989 0.323525    
## Pa_mm       -1.382e-05  1.205e-05  -1.147 0.252393    
## Latitude    -1.276e-03  2.826e-04  -4.516 9.07e-06 ***
## Longitude   -2.121e-05  8.975e-05  -0.236 0.813323    
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## 
## R-sq.(adj) =  0.271   Deviance explained = 28.3%
## GCV = 0.017125  Scale est. = 0.016787  n = 304
\end{verbatim}

\begin{verbatim}
## 
## Method: GCV   Optimizer: magic
## Model required no smoothing parameter selectionModel rank =  6 / 6
\end{verbatim}

\begin{verbatim}
## 
## Family: gaussian 
## Link function: identity 
## 
## Formula:
## log10(DeltaQf_perc/100 + 1) ~ DeltaF_perc + log10(Area_km2) + 
##     Pa_mm + Latitude + Longitude + From + length
## 
## Parametric coefficients:
##                   Estimate Std. Error t value Pr(>|t|)    
## (Intercept)     -1.027e+00  9.040e-01  -1.136 0.256902    
## DeltaF_perc     -1.267e-03  1.594e-04  -7.947 5.83e-14 ***
## log10(Area_km2) -2.003e-02  4.386e-03  -4.567 7.63e-06 ***
## Pa_mm           -7.871e-06  1.255e-05  -0.627 0.531281    
## Latitude        -1.084e-03  2.834e-04  -3.827 0.000163 ***
## Longitude        6.268e-05  9.560e-05   0.656 0.512595    
## From             5.612e-04  4.583e-04   1.225 0.221809    
## length          -2.931e-06  7.036e-06  -0.417 0.677294    
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## 
## R-sq.(adj) =  0.303   Deviance explained = 32.1%
## GCV = 0.016008  Scale est. = 0.01553   n = 268
\end{verbatim}

@@

\begin{verbatim}
## 
## Method: GCV   Optimizer: magic
## Smoothing parameter selection converged after 8 iterations.
## The RMS GCV score gradient at convergence was 2.042593e-08 .
## The Hessian was positive definite.
## Model rank =  21 / 21 
## 
## Basis dimension (k) checking results. Low p-value (k-index<1) may
## indicate that k is too low, especially if edf is close to k'.
## 
##               k'  edf k-index p-value    
## s(Area_km2) 2.00 1.00    0.97    0.29    
## s(Pa_mm)    2.00 1.25    0.78  <2e-16 ***
## s(From)     2.00 1.00    0.85    0.01 ** 
## s(length)   2.00 1.00    0.90    0.02 *  
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
\end{verbatim}

@@

\begin{verbatim}
## 
## Method: GCV   Optimizer: magic
## Smoothing parameter selection converged after 8 iterations.
## The RMS GCV score gradient at convergence was 2.042593e-08 .
## The Hessian was positive definite.
## Model rank =  21 / 21 
## 
## Basis dimension (k) checking results. Low p-value (k-index<1) may
## indicate that k is too low, especially if edf is close to k'.
## 
##               k'  edf k-index p-value    
## s(Area_km2) 2.00 1.00    0.97   0.330    
## s(Pa_mm)    2.00 1.25    0.78  <2e-16 ***
## s(From)     2.00 1.00    0.85   0.015 *  
## s(length)   2.00 1.00    0.90   0.095 .  
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
\end{verbatim}

@@

\begin{verbatim}
## 
## Method: GCV   Optimizer: magic
## Smoothing parameter selection converged after 5 iterations.
## The RMS GCV score gradient at convergence was 2.648312e-07 .
## The Hessian was positive definite.
## Model rank =  20 / 20 
## 
## Basis dimension (k) checking results. Low p-value (k-index<1) may
## indicate that k is too low, especially if edf is close to k'.
## 
##          k' edf k-index p-value    
## s(Pa_mm)  2   1     0.8  <2e-16 ***
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
\end{verbatim}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Zhang_all }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(DeltaF_perc, }\KeywordTok{log10}\NormalTok{(DeltaQf_perc}\OperatorTok{/}\DecValTok{100} \OperatorTok{+}\StringTok{ }\DecValTok{1}\NormalTok{), }\DataTypeTok{colour =}\NormalTok{ Assessment_technique,}\DataTypeTok{size =} \KeywordTok{log10}\NormalTok{(Area_km2))) }\OperatorTok{+}\StringTok{ }\KeywordTok{geom_point}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning: Removed 2 rows containing missing values (geom_point).
\end{verbatim}

\includegraphics{Forest_and_Water_files/figure-latex/unnamed-chunk-5-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{tiff}\NormalTok{(}\StringTok{"flow_forest_byArea.tiff"}\NormalTok{, }\DataTypeTok{width =} \DecValTok{2000}\NormalTok{, }\DataTypeTok{height =} \DecValTok{1400}\NormalTok{, }\DataTypeTok{res =} \DecValTok{300}\NormalTok{)}
\NormalTok{Zhang_all }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(DeltaF_perc}\OperatorTok{/}\DecValTok{100}\NormalTok{, (DeltaQf_perc}\OperatorTok{/}\DecValTok{100} \OperatorTok{+}\StringTok{ }\DecValTok{1}\NormalTok{), }\DataTypeTok{colour =}\NormalTok{ Assessment_technique, }\DataTypeTok{size =} \KeywordTok{log10}\NormalTok{(Area_km2))) }\OperatorTok{+}\StringTok{ }\KeywordTok{geom_point}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme_bw}\NormalTok{() }\OperatorTok{+}\StringTok{ }\KeywordTok{ylab}\NormalTok{(}\StringTok{"log10 (fractional change in flow + 1)"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{xlab}\NormalTok{(}\StringTok{"fractional change in forestry"}\NormalTok{) }\OperatorTok{+}\StringTok{ }\KeywordTok{scale_y_log10}\NormalTok{() }\OperatorTok{+}\StringTok{ }\KeywordTok{scale_size_continuous}\NormalTok{(}\DataTypeTok{name =} \StringTok{"log10 Area in km2"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{scale_colour_discrete}\NormalTok{(}\DataTypeTok{name =} \StringTok{"Assessment Technique"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning in self$trans$transform(x): NaNs produced
\end{verbatim}

\begin{verbatim}
## Warning: Transformation introduced infinite values in continuous y-axis
\end{verbatim}

\begin{verbatim}
## Warning: Removed 2 rows containing missing values (geom_point).
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{dev.off}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

@@

@@

\includegraphics{Forest_and_Water_files/figure-latex/unnamed-chunk-8-1.pdf}

@@

>>>>

In addition, additional variables added were the latitude and longitude
for the center of the watershed as an approximation of its spatial
location. Using this information annual average potential
evapotranspiration (E\_\{pot\}) was extracted from the MODIS16 ET
product, if a value of E0 was not available from the original papers.
This involved downloading the MODIS product for PET at 500 m scale for
the approximate catchment centroid using the package MODISTools via the
\href{https://modis.ornl.gov/data/modis_webservice.html}{MODIS/VIIRS
subsets} for the period 1 January 2000 - 31 December 2020. The average
annual PET calculated from this series was used as the an approximation
of average annual E0. For large watersheds, this value, similar to
annual average rainfall, is only an approximation of the climate at the
location.

The length of the study can be a variable influencing the change in flow
(e.g.~Jackson et al., 2005), as for example, more mature plantations are
thought to have smaller impacts on flow. Therefore, the length of the
study calculate as the difference between the starting data and
completion date of the different studies was extracted from the
references provided by Zhang et al. (2017).

@@

\textbf{This methods section is no longer correct and needs rewriting}

@@

\hypertarget{description-of-the-data}{%
\paragraph{description of the data}\label{description-of-the-data}}

The overall dataset contains 312 observations of changes in flow. The
overall distribution of changes in flow is highly skewed as is the
distribution of changes in forest cover and Area. The values of changes
in flow greater than 100\% and smaller than -100\% clearly create long
tails on the change in flow distribution. Note also the large number of
studies with 100\% forest cover reduction.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Zhang_all }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{pivot_longer}\NormalTok{(}\KeywordTok{c}\NormalTok{(DeltaF_perc, DeltaQf_perc, Area_km2, Pa_mm), }
               \DataTypeTok{names_to =} \StringTok{"variable"}\NormalTok{, }\DataTypeTok{values_to =} \StringTok{"value"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(value)) }\OperatorTok{+}\StringTok{ }\KeywordTok{geom_histogram}\NormalTok{(}\DataTypeTok{fill =} \StringTok{"steelblue"}\NormalTok{, }\DataTypeTok{bins=}\DecValTok{50}\NormalTok{) }\OperatorTok{+}\StringTok{ }\KeywordTok{theme_bw}\NormalTok{() }\OperatorTok{+}\StringTok{ }\KeywordTok{facet_wrap}\NormalTok{(}\OperatorTok{~}\NormalTok{variable, }\DataTypeTok{scales =} \StringTok{"free"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

The changes in forest cover contain both positive (forestation) and
negative values (deforestation). In ({\textbf{???}}) 2017, these changes
were analysed jointly, which assumes that the effect on the change in
flow is linear and non-hysteretic. However, it is a reasonable
hypothesis that this is not the case, and that the impact of an increase
in forest cover is different from the same fractional decrease in forest
cover. To be able to analyse this difference, all the change in forest
cover is converted to positive values, but an additional column is added
that indicates whether it was a forest cover increase or decrease.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Zhang_all2 <-}\StringTok{ }\NormalTok{Zhang_all }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{Forest_Sign =} \KeywordTok{ifelse}\NormalTok{(DeltaF_perc }\OperatorTok{<}\StringTok{ }\DecValTok{0}\NormalTok{,}
                              \StringTok{"decrease"}\NormalTok{, }\StringTok{"increase"}\NormalTok{),}
         \DataTypeTok{DeltaF_perc_pos =} \KeywordTok{ifelse}\NormalTok{(DeltaF_perc }\OperatorTok{<}\StringTok{ }\DecValTok{0}\NormalTok{,}
                                  \DecValTok{-1}\OperatorTok{*}\NormalTok{DeltaF_perc,}
\NormalTok{                                  DeltaF_perc))}
\end{Highlighting}
\end{Shaded}

\hypertarget{the-initial-relationship-between-change-in-forest-cover-and-streamflow}{%
\paragraph{The initial relationship between change in forest cover and
streamflow}\label{the-initial-relationship-between-change-in-forest-cover-and-streamflow}}

We start of with a simple linear regression (i.e.~following Zhang et al.
(2017)) just looking at the change in flow as a result in the percent
change forestry and including the direction of the change, increase in
forest cover, or decrease in forest cover.

\begin{verbatim}
## 
## Family: gaussian 
## Link function: identity 
## 
## Formula:
## DeltaQf_perc ~ DeltaF_perc_pos + Forest_Sign
## 
## Parametric coefficients:
##                     Estimate Std. Error t value Pr(>|t|)    
## (Intercept)           8.8466     6.5008   1.361    0.175    
## DeltaF_perc_pos       0.5364     0.1017   5.274 2.52e-07 ***
## Forest_Signincrease -32.4760     6.9367  -4.682 4.26e-06 ***
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## 
## R-sq.(adj) =   0.16   Deviance explained = 16.6%
## GCV = 3278.3  Scale est. = 3246.8    n = 312
\end{verbatim}

\includegraphics{Forest_and_Water_files/figure-latex/model1-1.pdf}

\begin{verbatim}
## 
## Method: GCV   Optimizer: magic
## Model required no smoothing parameter selectionModel rank =  3 / 3
\end{verbatim}

While the overall variance explained in this model is not high at 0.16,
it clearly indicates the hypothesised relationship between the change in
forest cover and the change in flow. The model suggests that for every
1\% change in forest cover, on the average, the flow changes 0.5\%.
However the change in flow is different for forest cover decreases
compared to forest cover increases. In fact, forest cover increases
decrease flow by 32\% less than a similar decrease in forest cover
causes flow to increase. So roughly speaking, a 1\% forest cover
increase on the average decreases flow by \((1 - 0.32)*0.5\%\), while a
the percentage forest cover decrease will increase flow by 0.5\%.

It is however clear from lack of explaining power, that there could be
confounding factors, as alluded to in the methods. The obvious ones
being catchment dryness and area (following Zhang et al. (2017)).

\begin{verbatim}
## 
## Family: gaussian 
## Link function: identity 
## 
## Formula:
## DeltaQf_perc ~ DeltaF_perc_pos + Forest_Sign + Area_km2 + Pa_mm
## 
## Parametric coefficients:
##                       Estimate Std. Error t value Pr(>|t|)    
## (Intercept)          2.323e+01  9.189e+00   2.528   0.0120 *  
## DeltaF_perc_pos      5.424e-01  1.030e-01   5.265 2.64e-07 ***
## Forest_Signincrease -3.324e+01  6.955e+00  -4.779 2.73e-06 ***
## Area_km2            -1.867e-06  1.322e-05  -0.141   0.8878    
## Pa_mm               -1.067e-02  4.944e-03  -2.158   0.0317 *  
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## 
## R-sq.(adj) =  0.168   Deviance explained = 17.9%
## GCV = 3269.3  Scale est. = 3217      n = 312
\end{verbatim}

\begin{verbatim}
## 
## Method: GCV   Optimizer: magic
## Model required no smoothing parameter selectionModel rank =  5 / 5
\end{verbatim}

Including area and annual precipitation does not really improve the
overall explaining power of the model, in fact, annual precipitation
appears to be only a very small confounding factor, representing only a
-0.01/\% partial effect in the change in streamflow, holding all other
factors constant. The catchment area does not appear to have an effect
at all in contrast to earlier reported studies (Filoso et al., 2017;
Zhang et al., 2017). The main effects remain the change in forest cover
and whether this is an increase or decrease

\hypertarget{the-effect-of-location-on-the-globe}{%
\paragraph{The effect of location on the
globe}\label{the-effect-of-location-on-the-globe}}

\begin{verbatim}
## 
## Family: gaussian 
## Link function: identity 
## 
## Formula:
## DeltaQf_perc ~ DeltaF_perc + Forest_Sign + Area_km2 + Pa_mm + 
##     Latitude + Longitude
## 
## Parametric coefficients:
##                       Estimate Std. Error t value Pr(>|t|)    
## (Intercept)          3.474e+01  1.055e+01   3.295  0.00110 ** 
## DeltaF_perc         -4.951e-01  1.064e-01  -4.652 4.95e-06 ***
## Forest_Signincrease -3.222e+00  1.160e+01  -0.278  0.78140    
## Area_km2            -9.488e-06  1.339e-05  -0.709  0.47903    
## Pa_mm               -1.007e-02  5.309e-03  -1.897  0.05883 .  
## Latitude            -3.806e-01  1.239e-01  -3.071  0.00233 ** 
## Longitude            1.074e-03  3.955e-02   0.027  0.97834    
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## 
## R-sq.(adj) =  0.172   Deviance explained = 18.8%
## GCV = 3336.1  Scale est. = 3259.8    n = 306
\end{verbatim}

\begin{verbatim}
## 
## Method: GCV   Optimizer: magic
## Model required no smoothing parameter selectionModel rank =  7 / 7
\end{verbatim}

@@

\begin{verbatim}
## 
## Family: gaussian 
## Link function: identity 
## 
## Formula:
## log10(DeltaQf_perc/100 + 1) ~ DeltaF_perc + log10(Area_km2) + 
##     Pa_mm + Latitude + Longitude + From + length
## 
## Parametric coefficients:
##                   Estimate Std. Error t value Pr(>|t|)    
## (Intercept)     -6.962e-01  1.110e+00  -0.627 0.531002    
## DeltaF_perc     -1.253e-03  1.612e-04  -7.772  1.8e-13 ***
## log10(Area_km2) -1.886e-02  5.186e-03  -3.637 0.000332 ***
## Pa_mm           -7.362e-06  1.257e-05  -0.586 0.558457    
## Latitude        -1.070e-03  2.823e-04  -3.791 0.000186 ***
## Longitude        6.832e-05  9.635e-05   0.709 0.478879    
## From             3.951e-04  5.604e-04   0.705 0.481473    
## length          -3.424e-04  7.026e-04  -0.487 0.626447    
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## 
## R-sq.(adj) =  0.303   Deviance explained = 32.1%
## GCV = 0.016004  Scale est. = 0.015526  n = 268
\end{verbatim}

@@

\begin{verbatim}
## 
## Method: GCV   Optimizer: magic
## Smoothing parameter selection converged after 8 iterations.
## The RMS GCV score gradient at convergence was 2.042593e-08 .
## The Hessian was positive definite.
## Model rank =  21 / 21 
## 
## Basis dimension (k) checking results. Low p-value (k-index<1) may
## indicate that k is too low, especially if edf is close to k'.
## 
##               k'  edf k-index p-value    
## s(Area_km2) 2.00 1.00    0.97   0.265    
## s(Pa_mm)    2.00 1.25    0.78  <2e-16 ***
## s(From)     2.00 1.00    0.85   0.005 ** 
## s(length)   2.00 1.00    0.90   0.060 .  
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
\end{verbatim}

@@

\begin{verbatim}
## 
## Method: GCV   Optimizer: magic
## Smoothing parameter selection converged after 8 iterations.
## The RMS GCV score gradient at convergence was 2.042593e-08 .
## The Hessian was positive definite.
## Model rank =  21 / 21 
## 
## Basis dimension (k) checking results. Low p-value (k-index<1) may
## indicate that k is too low, especially if edf is close to k'.
## 
##               k'  edf k-index p-value    
## s(Area_km2) 2.00 1.00    0.97   0.280    
## s(Pa_mm)    2.00 1.25    0.78  <2e-16 ***
## s(From)     2.00 1.00    0.85   0.015 *  
## s(length)   2.00 1.00    0.90   0.040 *  
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
\end{verbatim}

@@

\begin{verbatim}
## 
## Method: GCV   Optimizer: magic
## Smoothing parameter selection converged after 5 iterations.
## The RMS GCV score gradient at convergence was 2.648312e-07 .
## The Hessian was positive definite.
## Model rank =  20 / 20 
## 
## Basis dimension (k) checking results. Low p-value (k-index<1) may
## indicate that k is too low, especially if edf is close to k'.
## 
##          k' edf k-index p-value   
## s(Pa_mm)  2   1     0.8   0.005 **
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
\end{verbatim}

@@

\includegraphics{Forest_and_Water_files/figure-latex/unnamed-chunk-6-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Zhang_all2 }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(DeltaF_perc_pos, DeltaQf_perc, }\DataTypeTok{colour =}\NormalTok{ Assessment_technique,}\DataTypeTok{size =} \KeywordTok{log10}\NormalTok{(Area_km2), }\DataTypeTok{shape =}\NormalTok{ Forest_Sign)) }\OperatorTok{+}\StringTok{ }\KeywordTok{geom_point}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{tiff}\NormalTok{(}\StringTok{"flow_forest_byArea.tiff"}\NormalTok{, }\DataTypeTok{width =} \DecValTok{2500}\NormalTok{, }\DataTypeTok{height =} \DecValTok{1800}\NormalTok{, }\DataTypeTok{res =} \DecValTok{300}\NormalTok{)}
\NormalTok{Zhang_all2 }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(DeltaF_perc_pos, DeltaQf_perc, }\DataTypeTok{colour =}\NormalTok{ Assessment_technique,}\DataTypeTok{size =} \KeywordTok{log10}\NormalTok{(Area_km2), }
             \DataTypeTok{shape =}\NormalTok{ Forest_Sign)) }\OperatorTok{+}\StringTok{ }\KeywordTok{geom_point}\NormalTok{(}\DataTypeTok{alpha =} \FloatTok{0.5}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme_bw}\NormalTok{() }\OperatorTok{+}\StringTok{ }\KeywordTok{ylab}\NormalTok{(}\StringTok{"% change in flow"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{axis.title =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size =} \KeywordTok{rel}\NormalTok{(}\DecValTok{2}\NormalTok{)),}
        \DataTypeTok{axis.text =} \KeywordTok{element_text}\NormalTok{(}\DataTypeTok{size =} \KeywordTok{rel}\NormalTok{(}\FloatTok{1.5}\NormalTok{))) }\OperatorTok{+}
\StringTok{  }\KeywordTok{xlab}\NormalTok{(}\StringTok{"% change in forestry"}\NormalTok{) }\OperatorTok{+}\StringTok{ }\CommentTok{#scale_y_log10() +}
\StringTok{  }\KeywordTok{scale_size_continuous}\NormalTok{(}\DataTypeTok{name =} \StringTok{"log10(Area in km2)"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{scale_colour_discrete}\NormalTok{(}\DataTypeTok{name =} \StringTok{"Assessment Technique"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{scale_shape_discrete}\NormalTok{(}\DataTypeTok{name =} \StringTok{"Forest cover direction"}\NormalTok{)}
\KeywordTok{dev.off}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

@@

\includegraphics{Forest_and_Water_files/figure-latex/unnamed-chunk-9-1.pdf}

@@

\includegraphics{Forest_and_Water_files/figure-latex/unnamed-chunk-10-1.pdf}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{Zhang_all }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(length)) }\OperatorTok{+}\StringTok{ }\KeywordTok{geom_histogram}\NormalTok{(}\DataTypeTok{fill=}\StringTok{"blue"}\NormalTok{, }\DataTypeTok{bins =}\DecValTok{50}\NormalTok{) }
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning: Removed 42 rows containing non-finite values (stat_bin).
\end{verbatim}

\includegraphics{Forest_and_Water_files/figure-latex/unnamed-chunk-11-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{#windows()}
\NormalTok{Zhang_all }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\KeywordTok{log10}\NormalTok{(length), (DeltaQf_perc}\OperatorTok{/}\DecValTok{100} \OperatorTok{+}\StringTok{ }\DecValTok{1}\NormalTok{), }\DataTypeTok{colour =}\NormalTok{ DeltaF_perc}\OperatorTok{/}\DecValTok{100}\NormalTok{, }\DataTypeTok{size =} \KeywordTok{log10}\NormalTok{(Area_km2))) }\OperatorTok{+}\StringTok{ }\KeywordTok{geom_point}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme_bw}\NormalTok{() }\OperatorTok{+}\StringTok{ }\KeywordTok{ylab}\NormalTok{(}\StringTok{"log10 (fractional change in flow + 1)"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{xlab}\NormalTok{(}\StringTok{"Length of study"}\NormalTok{) }\OperatorTok{+}\StringTok{ }\KeywordTok{scale_size_continuous}\NormalTok{(}\DataTypeTok{name =} \StringTok{"log10 Area in km2"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{scale_colour_continuous}\NormalTok{(}\DataTypeTok{name =} \StringTok{"change in forest cover"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Warning: Removed 42 rows containing missing values (geom_point).
\end{verbatim}

\includegraphics{Forest_and_Water_files/figure-latex/unnamed-chunk-12-1.pdf}
